The Journey from Wild to Text Book Data: A Case Study of the National Longitudinal Survey of Youth

### 2021-04-07 report_alt.Rmd
Change the exploratory data analysis

In this part, we gave some examples of how the data might be used for data analysis in text book. The first example is that we observed relationship between wages and education. Using `features` function in `brolgar`, we extracted the median of the mean hourly age of each ID and plotted according to highest grade completed. We found in Figure \ref{fig:eda-plot} A that the higher the education completed does not necessarily means the higher wages earned. However, people who completed 12th grade tend to have higher wage compared to the other groups. Further, respondents who only completed 4th grade happen to have median about as high as the respondents who completed 10th grade.

@@

```{r, echo = FALSE}
wages_hs2020_tsibble <- wages_hs2020 %>%
  as_tsibble(key = id,
             index = year,
             regular = FALSE)

wages_hs2020_feature <- wages_hs2020_tsibble %>%
  features(mean_hourly_wage,
           feat_three_num)

wages_median_demog <- left_join(wages_hs2020_feature, demographic_nlsy79, by = "id") %>%
  mutate(hgc = factor(hgc, levels = c("1ST GRADE", "2ND GRADE", "3RD GRADE", 
                                      "4TH GRADE", "5TH GRADE", "6TH GRADE", 
                                      "7TH GRADE", "8TH GRADE", "9TH GRADE",
                                      "10TH GRADE", "11TH GRADE", "12TH GRADE", 
                                      "UNGRADED")))
```

```{r eda-plot, echo = FALSE, fig.height=6, fig.width=6, fig.cap="The example of EDA plot. Plot A shows the pattern of the mean hourly wage median against the highest grade completed. It shows that the higher eductaion completed does not necessarily means the more wages earned. Plot B and C show the mean hourly wage median over the years based on the race and gender respectively. Generally, the respondents experienced a positive trend over time despite of what their race and gender."}
hgc_plot <- ggplot(filter(wages_median_demog, hgc != "UNGRADED"), aes(y = hgc, x = med)) +
  geom_boxplot() +
  theme(axis.text.y = element_text(size = 6),
        axis.title = element_text(size = 9)) +
  ggtitle("A)") +
  xlab("median of mean hourly wage (US$)") +
  ylab("highest grade completed")

wages_hs2020_grouped <- wages_hs2020 %>%
  group_by(race, year) %>%
  summarise(median = median(mean_hourly_wage))

race_plot <- ggplot(wages_hs2020_grouped) +
  geom_line(aes(x = year,
                y = median,
                group = race,
                colour = race)) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 5),
        axis.title = element_text(size = 9)) +
  scale_color_brewer(palette = "Accent") +
  ggtitle("B)") +
  ylab("median of mean hourly wage (US$)")

wages_hs2020_grouped_gender <- wages_hs2020 %>%
  group_by(gender, year) %>%
  summarise(median = median(mean_hourly_wage))

gender_plot <- ggplot(wages_hs2020_grouped_gender) +
  geom_line(aes(x = year,
                y = median,
                group = gender,
                colour = gender)) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 5),
        axis.title = element_text(size = 9)) +
  scale_color_brewer(palette = "Dark2") +
  ggtitle("C)") +
  ylab("median of mean hourly wage (US$)")

hgc_plot / (race_plot + gender_plot)

#### using brolgar to explore the individual experiences <- change the EDA
```

>>>>

In this part, we gave some examples of how the data might be used for data analysis in text book. We used `brolgar` [@brolgar] to perform exploratory data analysis. In first example,

```{r wages-tsibble, echo = FALSE}
# create the wages data as tsibble
wages_hs2020_tsibble <- wages_hs2020 %>%
  as_tsibble(key = id,
             index = year,
             regular = FALSE)
```

```{r gender-plot, echo = FALSE}
# sample the data by gender

set.seed(31251587)
wages_hs2020_male <- wages_hs2020_tsibble %>%
  filter(gender == "MALE") %>%
  sample_n_keys(size = 120)

set.seed(31251587)
wages_hs2020_female <- wages_hs2020_tsibble %>%
  filter(gender == "FEMALE") %>%
  sample_n_keys(size = 120)

wages_hs2020_sample_gender <- bind_rows(wages_hs2020_male, wages_hs2020_female)

ggplot(wages_hs2020_sample_gender, 
       aes(x = year,
           y = mean_hourly_wage,
           group = id,
           color = gender)) + 
  geom_line(alpha = 0.4) +
  facet_strata(along = gender)
```

```{r race-plot, echo = FALSE}

set.seed(31251587)
wages_hs2020_white <- wages_hs2020_tsibble %>%
  filter(race == "NON-BLACK, NON-HISPANIC") %>%
  sample_n_keys(size = 60)

set.seed(31251587)
wages_hs2020_black <- wages_hs2020_tsibble %>%
  filter(race == "BLACK") %>%
  sample_n_keys(size = 60)

set.seed(31251587)
wages_hs2020_hisp <- wages_hs2020_tsibble %>%
  filter(race == "HISPANIC") %>%
  sample_n_keys(size = 60)

wages_hs2020_sample_race <- bind_rows(wages_hs2020_white, wages_hs2020_black, wages_hs2020_hisp)

ggplot(wages_hs2020_sample_race, 
       aes(x = year,
           y = mean_hourly_wage,
           group = id,
           color = race)) + 
  geom_line(alpha = 0.4) +
  facet_strata(along = race) +
  theme(legend.position = "bottom")
```

```{r wages-slope, echo = FALSE}
wages_slope <- key_slope(wages_hs2020_tsibble, 
                         mean_hourly_wage ~ year) %>%
  left_join(demographic_nlsy79, by = "id")

ggplot(wages_slope, aes(y = hgc,
                        x = .slope_year)) +
  geom_boxplot()
```

The first example is that we observed relationship between wages and education. Using `features` function in `brolgar`, we extracted the median of the mean hourly age of each ID and plotted according to highest grade completed. We found in Figure \ref{fig:eda-plot} A that the higher the education completed does not necessarily means the higher wages earned. However, people who completed 12th grade tend to have higher wage compared to the other groups. Further, respondents who only completed 4th grade happen to have median about as high as the respondents who completed 10th grade.

@@
